The Correspondence Analysis of Two Mode Networks Revisited

### 2025-03-30 analysis-R1.qmd
do it

```{r CA versus Bonacich Re-ordered Affiliation Matrices}
      # Affiliation matrix plot (CA ordering)
      A.ord <- as.matrix(A[order(CA.r$vectors[, 2]), order(CA.c$vectors[, 2])])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p1 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2))
      # Affiliation matrix plot (Bonacich ordering)
      A.ord <- as.matrix(A[order(eig.r[, 1]), order(eig.c[, 1])])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p2 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2)) 
     # Affiliation matrix plot (Bonacich ordering)
      A.ord <- as.matrix(A[order(eig.r[, 2]), order(eig.c[, 2])])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p3 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2)) 
      png(file = here("Plots", "ca-reord.png"), width=500, height=500)
      p1
      dev.off() 
      png(file = here("Plots", "bon-reord1.png"), width=500, height=500)
      p2
      dev.off() 
      png(file = here("Plots", "bon-reord2.png"), width=500, height=500)
      p3
      dev.off()
```

@@

```{r CA Eigenvector Plot} 
      eigvec.dat.r <- data.frame(rank = rank(CA.r$vectors[,2]), 
                             value = as.numeric(CA.r$vectors[,2]),
                             lab = names(CA.r$vectors[,2]))
      eigvec.dat.c <- data.frame(rank = rank(CA.c$vectors[,2]), 
                             value = as.numeric(CA.c$vectors[,2]),
                             lab = names(CA.c$vectors[,2]))
      eigval.dat.c <- data.frame(k = 1:13, 
                             value = CA.c$values[2:14]/max(CA.c$values[2:14])) %>% 
             mutate(value = value / max(value))
      # Eigenvector plot for the people
      p <- ggscatter(eigvec.dat.r, y = "rank", x = "value", point = FALSE,
                 font.label =  10, label = "lab", repel = TRUE)
      p <- p + theme(axis.text.y = element_blank(),
                 axis.line.x = element_blank(),
                 axis.line.y = element_blank(),
                 axis.ticks = element_blank(),
                 legend.position = "none")
      p <- p + labs(y = "Rank", x = "First Axis") 
      a <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")

@@

```{r CA Similarity and Correspondence Plot}
      norm.sim.r <- P.pp
      norm.sim.c <- P.gg
      diag(norm.sim.r) <- 0
      diag(norm.sim.c) <- 0
      norm.sim.r <- norm.sim.r[order(CA.r$vectors[,2]), 
                               order(CA.r$vectors[,2])]
      norm.sim.c <- norm.sim.c[order(CA.c$vectors[,2]), 
                               order(CA.c$vectors[,2])]

@@

      # Bonacich Correspondence Plot
      eig.d1 <- c(eig.r[, 1], eig.c[, 1])*-1
      eig.d2 <- c(eig.r[, 2], eig.c[, 2])*-1
      plot.dat <- data.frame(eig.d1, eig.d2, mode = c(rep(1, 18), rep(2, 14)))
      p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(mode), 
                                label = c(rownames(A), colnames(A))))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      p <- p + labs(x = "First Dimension", y = "Second Dimension")
      c <- p + scale_color_manual(values = c('red', 'blue'))
      png(file = here("Plots", "bon-corr-plot.png"), width=1500, height=1000)
      (a + b) / (c) + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off()
```

```{r CA versus SimRank}
      res <- SimRank(A)
      Sr <- res$Sr
      Sc <- res$Sc
      s1.r <- eigen(Sr)$vectors[, 2]
      s1.c <- eigen(Sc)$vectors[, 2]
      diag(Sr) <- 0
      diag(Sc) <- 0
      Sr <- Sr[order(d1.r), order(d1.r)]
      Sc <- Sc[order(d1.c), order(d1.c)]
      rownames(Sr) <- substr(colnames(P.pp), start = 1, stop = 3)
      rownames(Sc) <- substr(colnames(P.gg), start = 1, stop = 3)
      a <- ggcorrplot(Sr, lab = TRUE, show.diag = FALSE, lab_size = 3) +
         theme(legend.position = "none", 
               axis.text.x = element_text(hjust = -0.2)) + 
         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
                              low = "white", mid = "yellow", high = "purple")

      b <- ggcorrplot(Sc, lab = TRUE, show.diag = FALSE, lab_size = 3) +
         theme(legend.position = "none", 
               axis.text.x = element_text(hjust = -0.2)) + 
         scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
         scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5), 
                              low = "white", mid = "yellow", high = "purple")

      dat.r <- data.frame(ca = d1.r,
                          sr = s1.r,
                          lab = rownames(A))
      dat.c <- data.frame(ca = d1.c,
                          sr = s1.c,
                          lab = colnames(A))
      cor <- round(cor(dat.r[,c("ca", "sr")])[1, 2], 2)

      c <- ggscatter(dat.r, x = "ca", y = "sr", font.label= 8, 
                     label = "lab", repel = TRUE, 
                     add = "reg.line", 
                     add.params = list(color = "gray80", 
                                       linewidth = 0.5, linetype = 2)) +
                     annotate("text", x = -.2, y = .3, 
                              label = paste("r = ", cor), size = 6) 
      cor <- round(cor(dat.c[,c("ca", "sr")])[1, 2], 2)

      d <- ggscatter(dat.c, x = "ca", y = "sr", font.label= 8, 
                     label = "lab", repel = TRUE, 
                     add = "reg.line",
                     add.params = list(color = "gray80", 
                                       linewidth = 0.5, linetype = 2)) +
                     annotate("text", x = -.2, y = .3, 
                              label = paste("r = ", cor), size = 6) 
      png(file = here("Plots", "simrank-v-ca.png"), width=1000, height=900)
      (a + c + plot_layout(widths = c(2, 1))) / (b + d + plot_layout(widths = c(2, 1))) + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off() 
```

>>>>

@@

```{r CA Eigenvector Plot} 
      eigvec.dat.r <- data.frame(rank = rank(CA.r$vectors[,2]), 
                             value = as.numeric(CA.r$vectors[,2]),
                             lab = names(CA.r$vectors[,2]))
      eigvec.dat.c <- data.frame(rank = rank(CA.c$vectors[,2]), 
                             value = as.numeric(CA.c$vectors[,2]),
                             lab = names(CA.c$vectors[,2]))
      eigval.dat.c <- data.frame(k = 1:13, 
                             value = CA.c$values[2:14]/max(CA.c$values[2:14])) %>% 
             mutate(value = value / max(value))

      # Eigenvector plot for the people
      p <- ggscatter(eigvec.dat.r, y = "rank", x = "value", point = FALSE,
                 font.label =  10, label = "lab", repel = TRUE)
      p <- p + theme(axis.text.y = element_blank(),
                 axis.line.x = element_blank(),
                 axis.line.y = element_blank(),
                 axis.ticks = element_blank(),
                 legend.position = "none")
      p <- p + labs(y = "Rank", x = "First Axis") 
      a <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")

@@

```{r CA Similarity and Correspondence Plot}
      norm.sim.r <- P.pp
      norm.sim.c <- P.gg
      diag(norm.sim.r) <- 0
      diag(norm.sim.c) <- 0
      norm.sim.r <- norm.sim.r[order(CA.r$vectors[,2]), 
                               order(CA.r$vectors[,2])]
      norm.sim.c <- norm.sim.c[order(CA.c$vectors[,2]), 
                               order(CA.c$vectors[,2])]
      # CA Spectral Clustering
      ca.dat <- data.frame(rbind(CA.r$vectors[,2:4], CA.c$vectors[,2:4]))
      ca.clus.p <- kmeans(ca.dat[1:18, ], centers = 3, nstart = 25)
      ca.clus.g <- kmeans(ca.dat[19:32, ], centers = 3, nstart = 25)
      ca.clus <- kmeans(ca.dat, centers = 3, nstart = 25)

@@

      # K-means clustering of Bonacich solution
      eig.dat <- data.frame(d1 = eig.d1, d2 = eig.d2)
      eig.clus.p <- kmeans(eig.dat[1:18,], centers = 3, nstart = 25)$cluster
      eig.clus.g <- kmeans(eig.dat[19:32,], centers = 3, nstart = 25)$cluster

      # Bonacich Correspondence Plot
      eig.d1 <- c(eig.r[, 1], eig.c[, 1])*-1
      eig.d2 <- c(eig.r[, 2], eig.c[, 2])*-1
      plot.dat <- data.frame(eig.d1, eig.d2, 
                             mode = c(rep(1, 18), rep(2, 14)),
                             clus = c(eig.clus.p, eig.clus.g+3)
                             )
      p <- ggplot(plot.dat, aes(y = eig.d2, x = eig.d1, color = factor(clus), 
                                label = c(rownames(A), colnames(A))))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50, fontface = "bold")
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      c <- p + labs(x = "First Dimension", y = "Second Dimension")
      #c <- p + scale_color_manual(values = c('red', 'blue'))
      png(file = here("Plots", "bon-corr-plot.png"), width=1500, height=1000)
      (a + b) / (c) + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off()
```

```{r SimRank Similarity Matrices}
   SR.res <- SimRank(A)
   SR.p <- SR.res$Sr
   SR.g <- SR.res$Sc
   rownames(SR.p) <- substr(colnames(SR.p), start = 1, stop = 3)
   rownames(SR.g) <- substr(colnames(SR.g), start = 1, stop = 3)
   a <- ggcorrplot(SR.p, lab = TRUE, lab_size = 5,
                     colors = c(rep("white", 3))) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2, size = 16),
            axis.text.y = element_text(size = 16)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
   b <- ggcorrplot(SR.g, lab = TRUE, lab_size = 5,
                     colors = c(rep("white", 3))) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2, size = 16),
            axis.text.y = element_text(size = 16)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
   png(file = here("Plots", "p-SimRank.png"), width=900, height=900)
   P.p
   dev.off() 
   png(file = here("Plots", "g-SimRank.png"), width=900, height=900)
   P.g
   dev.off() 
```

```{r CA versus SimRank}
      # SimRank Eigenvector plot data
      SR.eig.p <- eigen(SR.res$Sr)$vectors
      SR.eig.g <- eigen(SR.res$Sc)$vectors
      eigvec.dat.r <- data.frame(rank = rank(SR.eig.p[, 2]), 
                             value = as.numeric(SR.eig.p[, 2]),
                             lab = rownames(A))
      eigvec.dat.c <- data.frame(rank = rank(SR.eig.g[, 2]), 
                             value = as.numeric(SR.eig.g[, 2]),
                             lab = colnames(A))

      # Eigenvector plot for the people
      p <- ggscatter(eigvec.dat.r, y = "rank", x = "value", point = FALSE,
                 font.label =  10, label = "lab", repel = TRUE)
      p <- p + theme(axis.text.y = element_blank(),
                 axis.line.x = element_blank(),
                 axis.line.y = element_blank(),
                 axis.ticks = element_blank(),
                 legend.position = "none")
      p <- p + labs(y = "Rank", x = "First Axis") 
      a <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")

      # Data for scatterplot
      dat.p <- data.frame(ca = CA.r$vectors[, 2],
                          sr = SR.eig.p[, 2],
                          lab = rownames(A))
      dat.g <- data.frame(ca = CA.c$vectors[, 2],
                          sr = SR.eig.g[, 2],
                          lab = colnames(A))

      # SimRank versus CA (People)
      cor <- round(cor(dat.p[,c("ca", "sr")])[1, 2], 2)
      c <- ggscatter(dat.p, x = "ca", y = "sr", font.label= 8, 
                     label = "lab", repel = TRUE, 
                     add = "reg.line", 
                     add.params = list(color = "gray80", 
                                       linewidth = 0.5, linetype = 2)) +
                     annotate("text", x = -.2, y = .3, 
                              label = paste("r = ", cor), size = 6) +
                     labs(x = "CA Scores", y = "SimRank Scores")

      # SimRank versus CA (Groups)
      cor <- round(cor(dat.g[,c("ca", "sr")])[1, 2], 2)
      d <- ggscatter(dat.g, x = "ca", y = "sr", font.label= 8, 
                     label = "lab", repel = TRUE, 
                     add = "reg.line",
                     add.params = list(color = "gray80", 
                                       linewidth = 0.5, linetype = 2)) +
                     annotate("text", x = -.2, y = .3, 
                              label = paste("r = ", cor), size = 6) +
                     labs(x = "CA Scores", y = "SimRank Scores")

      # Composite Plot
      png(file = here("Plots", "simrank-v-ca.png"), width=1000, height=900)
         (a + b)/(c + d) + plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off() 
```

```{r Blocked Affiliation Matrices}
      # First CA Dimension
      A.ord <- as.matrix(A[order(CA.r$vectors[, 2]), order(CA.c$vectors[, 2])])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p1 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2))
      # First Bonacich Dimension
      A.ord <- as.matrix(A[order(eig.r[, 1]), order(eig.c[, 1])])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p2 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2)) 
     # Second Bonacich Dimension
      A.ord <- as.matrix(A[order(eig.r[, 2]), order(eig.c[, 2])])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p3 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2)) 
     # Second SimRank Eigenvector
      A.ord <- as.matrix(A[order(s2.r), order(s2.c)])
      p <- ggcorrplot(t(A.ord)) 
      p <- p + scale_x_discrete(position = "top")
      p <- p + scale_y_discrete(limits = rev)
      p4 <- p + theme(legend.position = "none",
                               axis.text.x = element_text(hjust = -0.2))
      # Saving plots
      png(file = here("Plots", "ca-reord1.png"), width=500, height=500)
      p1
      dev.off() 
      png(file = here("Plots", "bon-reord1.png"), width=500, height=500)
      p2
      dev.off() 
      png(file = here("Plots", "bon-reord2.png"), width=500, height=500)
      p3
      dev.off()
      png(file = here("Plots", "SR-reord1.png"), width=500, height=500)
      p4
      dev.off()
```
